---
title: "Statistical Inference - Quiz 02"
author: "Utkarsh Ashok Pathrabe"
output: md_document
---

Quiz 02
=======  

|Attempts|Score|  
|:------:|:---:|  
|   1/3  | 8/8 |  

Question 01
-----------  
What is the variance of the distribution of the average an IID draw of n observations from a population with mean $\mu$ and variance $\sigma^2$.  

### Answer  
$\sigma^2$/n  

Question 02
-----------  
Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?   

### Answer  
* 16%  

#### Explanation  
```{r warning=FALSE, error=FALSE}
pnorm(70, mean=80, sd=10)
```  

Question 03  
-----------  
Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?  

### Answer  
* Approximately 1223.  

#### Explanation  
```{r warning=FALSE, error=FALSE}
qnorm(.95, mean=1100, sd=75)
```  

Question 04
-----------  
Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?  

### Answer  
* Approximately 1112 cc.  

#### Explanation  
```{r warning=FALSE, error=FALSE}
qnorm(.95, mean=1100, sd=(75/sqrt(100)))
```  

Question 05
-----------  
You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?  

### Answer  
* 19%  

#### Explanation  
```{r warning=FALSE, error=FALSE}
1 - pbinom(3, 5, .5)
```  

Question 06
-----------  
The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?  

### Answer  
* 68%  

#### Explanation  
```{r warning=FALSE, error=FALSE}
sd <- 10 / sqrt(100)
pnorm(16, mean=15, sd) - pnorm(14, mean=15, sd)
```  

Question 07
-----------  
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?  

### Answer  
* 0.5  

Question 08
-----------  
The number of people showing up at a bus stop is assumed to be Poisson with a mean of 5 people per hour. You watch the bus stop for 3 hours. About what's the probability of viewing 10 or fewer people?  

### Answer  
* 0.12  

#### Explanation  
```{r warning=FALSE, error=FALSE}
ppois(10, 3*5)
```  