---
title: "Statistical Inference Project Part 1 - A Simulation Exercise"
author: "Utkarsh Ashok Pathrabe"
output: pdf_document
---

Overview:  
---------  
This project investigates the Exponential distribution in R and compares it with the Central Limit Theorem. The mean of the Exponential distribution is $\frac{1}{\lambda}$ and the standard deviation is also $\frac{1}{\lambda}$. A thousand simulations of the distribution of 40 exponentials would be investigated.  

Simulations:  
------------  
The exponential distribution can be simulated in R with `rexp(n, lambda)`, where `lambda` is the rate parameter and `n` is the number of observations. For the purpose of all the simulations in this project, value of `lambda` is set to `0.2`.  
First we load the `ggplot2` plotting library.  
```{r warning=FALSE, error=FALSE}
library(ggplot2)
```  
We then initialize the simulation controlling variables.  
```{r warning=FALSE, error=FALSE}
noSim <- 1000
sampSize <- 40
lambda <- 0.2
```  
Set the seed of the Random Number Generator, so that the analysis is reproducible.  
```{r warning=FALSE, error=FALSE}
set.seed(3)
```  
Create a matrix with thousand rows corresponding to 1000 simulations and forty columns corresponding to each of 40 random simulations.  
```{r warning=FALSE, error=FALSE}
simulationMatrix <- matrix(rexp(n = noSim * sampSize, rate = lambda), noSim, sampSize)
```  
Create a vector of thousand rows containing the mean of each row of the `simulationMatrix`.  
```{r warning=FALSE, error=FALSE}
simulationMean <- rowMeans(simulationMatrix)
```  
Create a data frame containing the whole data.  
```{r warning=FALSE, error=FALSE}
simulationData <- data.frame(cbind(simulationMatrix, simulationMean))
```  
We plot the simulation data to visualize it.  
```{r warning=FALSE, error=FALSE}
ggplot(data = simulationData, aes(simulationData$simulationMean)) + 
  geom_histogram(breaks = seq(2, 9, by = 0.2), col = "blue", aes(fill = ..count..)) + 
  labs(title = "Histogram of Mean Distribution", x = "Simulation Means", y = "Frequency") + 
  geom_vline(aes(xintercept=mean(simulationData$simulationMean)), color="red", 
             linetype="dashed", size=1)
```  

Sample Mean Versus Theoretical Mean:  
------------------------------------  
The actual mean of the simulated mean sample data is `r mean(simulationMean)`, calculated by:  
```{r warning=FALSE, error=FALSE}
actualMean <- mean(simulationMean)
```  
And the theoretical mean is `r 1/lambda`, calculated by:  
```{r warning=FALSE, error=FALSE}
theoreticalMean <- (1 / lambda)
```  
Thus, we can see that the actual mean of the simulated mean sample data is very close to the theoretical mean of original data distribution.  

Sample Variance Versus Theoretical Variance:  
--------------------------------------------  
The actual variance of the simulated mean sample data is `r var(simulationMean)`, calculated by:  
```{r warning=FALSE, error=FALSE}
actualVariance <- var(simulationMean)
```  
And the theoretical variance is `r ((1/lambda)^2)/sampSize`, calculated by:  
```{r warning=FALSE, error=FALSE}
theoreticalVariance <- ((1 / lambda) ^ 2) / sampSize
```  
Thus, we can see that the actual variance of the simulated mean sample data is very close to the theoretical variance of original data distribution.  

Distribution:  
-------------  
To prove that the simulated mean sample data approximately follows the Normal distribution, we perform the following three steps:  

#### Step 1: Create an approximate normal distribution and see how the sample data alligns with it.  
```{r warning=FALSE, error=FALSE}
qplot(simulationMean, geom = 'blank') + 
  geom_line(aes(y=..density.., colour='Empirical'), stat='density', size=1) + 
  stat_function(fun=dnorm, args=list(mean=(1/lambda), sd=((1/lambda)/sqrt(sampSize))), 
                aes(colour='Normal'), size=1) + 
  geom_histogram(aes(y=..density.., fill=..density..), alpha=0.4, 
                 breaks = seq(2, 9, by = 0.2), col='red') + 
  scale_fill_gradient("Density", low = "yellow", high = "red") + 
  scale_color_manual(name='Density', values=c('brown', 'blue')) + 
  theme(legend.position = c(0.85, 0.60)) + 
  labs(title = "Mean Density Distribution", x = "Simulation Means", y = "Density")
```  

From above histogram, the simulated mean sample data can be adequately approximated with the normal distribution.  

#### Step 2: Compare the 95% confidence intervals of the simulated mean sample data and the theoretical normally distributed data.  
```{r warning=FALSE, error=FALSE}
actualConfInterval <- actualMean+c(-1,1)*1.96*sqrt(actualVariance)/sqrt(sampSize)
theoreticalConfInterval <- theoreticalMean+c(-1,1)*1.96*
  sqrt(theoreticalVariance)/sqrt(sampSize)
```  
Actual 95% confidence interval is [`r actualConfInterval[1]`, `r actualConfInterval[2]`] and Theoretical 95% confidence interval is [`r theoreticalConfInterval[1]`, `r theoreticalConfInterval[2]`] and we see that both of them are approximately same.  

#### Step 3: q-q Plot for Qunatiles.  
```{r warning=FALSE, error=FALSE}
qqnorm(simulationMean)
qqline(simulationMean)
```  

The actual quantiles also closely match the theoretical quantiles, hence the above three steps prove that the distribution is approximately normal.  